Individual heterogeneity in the functional network topography of medial prefrontal

### 2019-09-09 Manuscript.Rmd
formatting tweaks

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are implicated in multiple cognitive functions, including subjective valuation processes and processes linked to the default network (DN). Our ability to interpret these seemingly co-localized effects is constrained by a limited understanding of the individual-level heterogeneity in mPFC/PCC functional organization. Here we used cortical surface-based meta-analysis to identify a parcel in human PCC that was preferentially implicated in DN effects relative to valuation. We then used resting-state fMRI data and a data-driven network analysis algorithm, spectral partitioning, to partition mPFC and PCC into "DN" and "non-DN" subdivisions in individual participants (n = 100 from the Human Connectome Project). The spectral partitioning algorithm efficiently identified individual-level cortical subdivisions that were reliable across test/retest sessions and varied markedly across individuals, especially in mPFC. Our results point toward a new generation of strategies for assessing whether distinct cognitive functions engage common or distinct mPFC subregions.

## Introduction

@@

## Materials and Methods

### *Code Accessibility*

@@

### *Search space*

@@

### *Meta-analysis*

@@

### *Resting-state fMRI Data*

@@

### *Network Definition*

@@

### *Community Detection*

@@

### *Partition Evaluation*

@@

### *Seed-based Resting-state Functional Connectivity versus Community Detection*

@@

### *Associations with sulcal morphology*

@@

## Results

### *Meta-analysis*

@@

### *Individual-level DN and non-DN communities*

@@

### *Pattern variability over time*

@@

### *Test/re-test reliability across days*

@@

### *Test/re-test reliability across runs*

@@

### *Correlation vs community detection in mPFC*

@@

### *Relationship between functional organization and sulcal morphology*

@@

Next, we asked if the idiosyncratic organization of the DN was reflected by patterns of sulcal morphology. A variety of studies have provided evidence that suclal and gyral organization inform the location of functional effects [@Amiez2013; @Amiez2014; @Zlatkina2016]. Among these, recent work has suggested that the location of the DN in individuals lies mostly within sulci in vmPFC [@Lopez2019]. We sought to reproduce this relationship using the SP communities, as well as interrogating whether it persisted in PCC and more dorsal mPFC regions.

@@

We quantified these observations by dividing the search space into three regions: a ventral mPFC area corresponding to the one used by Lopez-Persem et al. [-@Lopez2019], a superior mPFC zone (sup-mPFC) that contained the remaining dorsal mPFC regions, and PCC (Figure 8B, left). For each individual, we correlated the unthresholded Fiedler vector with cortical curvature for each combination of region and hemisphere. Negative correlations in this context imply that the DN was likely found in sulci and non-DN in gyri. Results can be found on Figure 8B. We found that these features tended to be negatively correlated in both left (mean correlation = `r round(curvStats$meanCor[5], digits = 2)`, SE = `r round(curvStats$seCor[5], digits = 2)`) and right (mean correlation = `r round(curvStats$meanCor[6], digits = 2)`, SE = `r round(curvStats$seCor[6], digits = 2)`) hemispheres of vmPFC across individuals, but this distribution was only significantly different than zero on the right hemisphere (one sample t-test: t = `r curvStats$TStat[6]`, *p* `r ifelse(curvStats$p[6] < 0.001, "< 0.001", paste("=", curvStats$p[6]))`; Cohen's D = `r round(curvStats$ES[6], digits = 2)`). Correlations between FV and curvature were positive and significantly different than zero in both left (mean correlation = `r round(curvStats$meanCor[3], digits = 2)`, SE = `r round(curvStats$seCor[3], digits = 2)`; one sample t-test: t = `r curvStats$TStat[3]`, *p* `r ifelse(curvStats$p[3] < 0.001, "< 0.001", paste("=", curvStats$p[3]))`; Cohen's D = `r round(curvStats$ES[3], digits = 2)`) and right (mean correlation = `r round(curvStats$meanCor[4], digits = 2)`, SE = `r round(curvStats$seCor[4], digits = 2)`; one sample t-test: t = `r curvStats$TStat[4]`, *p* `r ifelse(curvStats$p[4] < 0.001, "< 0.001", paste("=", curvStats$p[4]))`; Cohen's D = `r round(curvStats$ES[4], digits = 2)`) hemispheres of sup-mPFC. Unlike frontal regions, correlations in PCC were negative and significantly different than zero in the left hemisphere (mean correlation = `r round(curvStats$meanCor[1], digits = 2)`, SE = `r round(curvStats$seCor[1], digits = 2)`; one sample t-test: t = `r curvStats$TStat[1]`, *p* `r ifelse(curvStats$p[1] < 0.001, "< 0.001", paste("=", curvStats$p[1]))`; Cohen's D = `r round(curvStats$ES[1], digits = 2)`), but significantly above zero in the right hemisphere (mean correlation = `r round(curvStats$meanCor[2], digits = 2)`, SE = `r round(curvStats$seCor[2], digits = 2)`; one sample t-test: t = `r curvStats$TStat[2]`, *p* `r ifelse(curvStats$p[2] < 0.001, "< 0.001", paste("=", curvStats$p[2]))`; Cohen's D = `r round(curvStats$ES[2], digits = 2)`). The difference across hemispheres was significant (paired t-test: t = `r round(curvPCC_t$statistic, digits = 2)`, *p* `r ifelse(curvPCC_t$p.val < 0.001, "< 0.001", paste("=", curvPCC_t$p.val))`; Cohen's D = `r round(curvPCC_ES, digits = 2)`). These results are in line with what can be observed in individual brains, partially replicate the findings from Lopez-Persem et al. [-@Lopez2019], and suggest that the association between function and structure is heterogenous within canonical DN.

### *Alignment of mPFC community structure with a proposed DN sub-network organization*

@@

## Discussion

@@

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. The precise spatial positioning of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individual. The idiosyncrasy in functional topography was particularly pronounced in mPFC, and was identified in both run-based and day-based analyses. Individual-specificity could theoretically arise from a variety of sources. For example, Individual variability could be due to shifts in functional organization that are independent of structural features [@Haxby2013; @Nenning2017], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. Alternatively, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to vary systematically across individuals [@Mackey2014]. Our findings support this idea, echoing previous findings that DN is contained within sulci (in particular the superior rostral sulcus) in vmPFC [@Lopez2019]. This relationship was heterogeneous throughout the rest of the search space, as more dorsal mPFC regions were likely located in gyri, whereas PCC had hemispheric differences in this association.  An important goal for future work will be to assess whether the network layout in these regions can also be predicted on the basis of other aspects of brain structure, such as myeloarchitecture [@Glasser2016] or structural connectivity [@Osher2016; @Saygin2011; @Saygin2016].

@@

>>>>

## Highights:

- Medial Default Network overlaps topographically with task effects such as valuation.

- Qualitative work suggests overlap could be due to variability across individuals.

- We used Spectral Partitioning to find and compare Default Network across individuals.

- Topographical organization was highly idiosyncratic, especially in frontal regions.

- Variability suggests that apparent overlap should be re-examined in individuals.

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are implicated in multiple cognitive functions, including subjective valuation processes and processes linked to the default network (DN). Our ability to interpret these seemingly co-localized effects is constrained by a limited understanding of the individual-level heterogeneity in mPFC/PCC functional organization. Here we used cortical surface-based meta-analysis to identify a parcel in human PCC that was preferentially implicated in DN effects relative to valuation. We then used resting-state fMRI data and a data-driven network analysis algorithm, spectral partitioning, to partition mPFC and PCC into "DN" and "non-DN" subdivisions in individual participants (n = 100 from the Human Connectome Project). The spectral partitioning algorithm identified individual-level cortical subdivisions that were reliable across test/retest datasets and varied markedly across individuals, especially in mPFC. Location of the DN was informed by cortical folding, and resembled a recently proposed fractionation of this network. Our results point toward a new generation of strategies for assessing whether distinct cognitive functions engage common or distinct mPFC subregions.

@@

## 1. Introduction

@@

## 2. Material and Methods

### 2.1. *Code Accessibility*

@@

### 2.2. *Search space*

@@

### 2.3. *Meta-analysis*

@@

### 2.4. *Resting-state fMRI Data*

@@

### 2.5. *Network Definition*

@@

### 2.6. *Community Detection*

@@

### 2.7. *Partition Evaluation*

@@

### 2.8. *Seed-based Resting-state Functional Connectivity versus Community Detection*

@@

### 2.9. *Associations with sulcal morphology*

@@

## 3. Results

### 3.1. *Meta-analysis*

@@

### 3.2. *Individual-level DN and non-DN communities*

@@

### 3.3. *Pattern variability over time*

@@

### 3.4. *Test/re-test reliability across days*

@@

### 3.5. *Test/re-test reliability across runs*

@@

### 3.6. *Correlation vs community detection in mPFC*

@@

### 3.7. *Relationship between functional organization and sulcal morphology*

@@

Next, we asked if the idiosyncratic organization of the DN was reflected by patterns of sulcal morphology. A variety of studies have provided evidence that suclal and gyral organization inform the location of functional effects [@Amiez2013; @Amiez2014; @Zlatkina2016]. Among these, recent work has suggested that the location of the DN in individuals lies mostly within sulci in vmPFC [@Lopez2019]. We sought to reproduce this relationship using the SP communities, and interrogated whether it persisted in PCC and more dorsal mPFC regions.

@@

We quantified these observations by dividing the search space into three regions: a ventral mPFC area corresponding to the one used by Lopez-Persem et al. [-@Lopez2019], a superior mPFC zone (sup-mPFC) that contained the remaining dorsal mPFC regions, and PCC (Figure 8B, left). For each individual, we correlated the unthresholded Fiedler vector with cortical curvature for each combination of region and hemisphere. Negative correlations in this context imply that the DN was likely found in sulci and non-DN in gyri. Results can be found on Figure 8B. We found that these features tended to be negatively correlated in both left (mean correlation = `r round(curvStats$meanCor[5], digits = 2)`, SE = `r round(curvStats$seCor[5], digits = 2)`) and right (mean correlation = `r round(curvStats$meanCor[6], digits = 2)`, SE = `r round(curvStats$seCor[6], digits = 2)`) hemispheres of vmPFC across individuals, but this distribution was only significantly different than zero on the right hemisphere (one sample t-test: t = `r curvStats$TStat[6]`, *p* `r ifelse(curvStats$p[6] < 0.001, "< 0.001", paste("=", curvStats$p[6]))`; Cohen's D = `r round(curvStats$ES[6], digits = 2)`). Correlations between FV and curvature were positive and significantly different from zero in both left (mean correlation = `r round(curvStats$meanCor[3], digits = 2)`, SE = `r round(curvStats$seCor[3], digits = 2)`; one sample t-test: t = `r curvStats$TStat[3]`, *p* `r ifelse(curvStats$p[3] < 0.001, "< 0.001", paste("=", curvStats$p[3]))`; Cohen's D = `r round(curvStats$ES[3], digits = 2)`) and right (mean correlation = `r round(curvStats$meanCor[4], digits = 2)`, SE = `r round(curvStats$seCor[4], digits = 2)`; one sample t-test: t = `r curvStats$TStat[4]`, *p* `r ifelse(curvStats$p[4] < 0.001, "< 0.001", paste("=", curvStats$p[4]))`; Cohen's D = `r round(curvStats$ES[4], digits = 2)`) hemispheres of sup-mPFC. Unlike frontal regions, correlations in PCC were negative and significantly different than zero in the left hemisphere (mean correlation = `r round(curvStats$meanCor[1], digits = 2)`, SE = `r round(curvStats$seCor[1], digits = 2)`; one sample t-test: t = `r curvStats$TStat[1]`, *p* `r ifelse(curvStats$p[1] < 0.001, "< 0.001", paste("=", curvStats$p[1]))`; Cohen's D = `r round(curvStats$ES[1], digits = 2)`), but significantly above zero in the right hemisphere (mean correlation = `r round(curvStats$meanCor[2], digits = 2)`, SE = `r round(curvStats$seCor[2], digits = 2)`; one sample t-test: t = `r curvStats$TStat[2]`, *p* `r ifelse(curvStats$p[2] < 0.001, "< 0.001", paste("=", curvStats$p[2]))`; Cohen's D = `r round(curvStats$ES[2], digits = 2)`). The difference across hemispheres was significant (paired t-test: t = `r round(curvPCC_t$statistic, digits = 2)`, *p* `r ifelse(curvPCC_t$p.val < 0.001, "< 0.001", paste("=", curvPCC_t$p.val))`; Cohen's D = `r round(curvPCC_ES, digits = 2)`). These results are in line with what can be observed in individual brains, partially replicate the findings from Lopez-Persem et al. [-@Lopez2019], and suggest that the association between function and structure is heterogenous within canonical DN.

### 3.8. *Alignment of mPFC community structure with a proposed DN sub-network organization*

@@

## 4. Discussion

@@

We found a generalizable pattern across individual partitionings, in which the DN community covered ventral/dorsal mPFC and posterior PCC, with the non-DN community concentrated in pregenual ACC and anterior PCC. The precise spatial positioning of this general community structure was highly heterogenous across individuals, yet stable across test/re-test evaluations within-individual. The idiosyncrasy in functional topography was particularly pronounced in mPFC, and was identified in both run-based and day-based analyses. Individual-specificity could theoretically arise from a variety of sources. For example, Individual variability could be due to shifts in functional organization that are independent of structural features [@Haxby2013; @Nenning2017], or be characterized by the pattern of functional connections with the rest of the brain [@Mars2018; @Passingham2002; @Tobyne2018]. Alternatively, the functional topography of mPFC could be governed by its underlying sulcal and gyral organization, which has been shown to vary systematically across individuals [@Mackey2014]. Our results support this idea, echoing previous findings that DN is contained within sulci (in particular the superior rostral sulcus) in vmPFC [@Lopez2019]. This relationship was heterogeneous throughout the rest of the search space. DN was likely located in gyri in more dorsal mPFC regions, whereas this association differed across hemispheres in PCC. An important goal for future work will be to assess whether the network layout in these regions can also be predicted on the basis of other aspects of brain structure, such as myeloarchitecture [@Glasser2016] or structural connectivity [@Osher2016; @Saygin2011; @Saygin2016].

@@

### 4.1. *Conclusion*
